---
title: "DSPI Model Markdown"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_width: 9
    fig_height: 7
---

Script for mixed model analysis of Diurnal Saturated Photosynthesis Index DSPI
This is a two species analysis and needs to be run in sequence. You can skip
the first species but will need to run some of the code located in the Ulva section
in order to run linear regressions with growth.

Load the various libraries
```{r, message = FALSE}
library(lme4)
library(lmerTest)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
library(rstatix)
#for plots and tables
library(ggplot2)
library(ggpubr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(sjPlot)
library(sjmisc)
library(mmtable2)
library(gt)
library(purrr)
library(stringr)
library(tidyr)
```

Load the dataset
```{r}
ek_irrad_data <- read.csv("input_data/mean_supersaturation_by_period.csv")
```

Assign several variables as factors
```{r}
ek_irrad_data$Run <- as.factor(ek_irrad_data$Run)
ek_irrad_data$Temperature <- as.factor(ek_irrad_data$Temp...C.)
ek_irrad_data$Treatment <- as.factor(as.character(ek_irrad_data$Treatment))
ek_irrad_data$RLC.Order <- as.factor(ek_irrad_data$RLC.Order)
```

Convert carbon_total from micromol to mol
```{r}
ek_irrad_data$carbon_mol <- ek_irrad_data$carbon_total / 1e+6
```

Subset data by species, remove treatment 2.5 for Ulva
```{r}
ulva <- subset(ek_irrad_data, Species == "ul" & Treatment != 2.5)
ulva$treatment_graph[ulva$Treatment == 0] <- "1) 35ppt/0.5umol"
ulva$treatment_graph[ulva$Treatment == 1] <- "2) 35ppt/14umol" 
ulva$treatment_graph[ulva$Treatment == 2] <- "3) 28ppt/27umol" 
ulva$treatment_graph[ulva$Treatment == 3] <- "5) 18ppt/53umol" 
ulva$treatment_graph[ulva$Treatment == 4] <- "6) 11ppt/80umol"
ulva$treatment_graph[ulva$Treatment == 2.5] <- "4) 28ppt/53umol"
```

ULVA____________________________________________________________
 
Make a histogram and residual plots of the data for ulva
```{r}
hist(ulva$carbon_mol, main = paste("Ulva lactuca"), col = "olivedrab3", labels = TRUE)

ulva %>% ggplot(aes(carbon_mol)) +
        geom_histogram(binwidth=0.25, fill = "#5BB300", color = "black", size = 0.25, alpha = 0.85) +
        theme_bw()
```
Run model without interaction between the treatments and temperature - supersat_avg is in minutes
Take RLC.Order our of the random effects because causing problems of singularity. R2 is same with or without (+ (1 | RLC.Order))
```{r}
dspi_model_ulva <- lmer(formula = carbon_mol ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva)
```

Plot residuals
```{r}
hist(resid(dspi_model_ulva))
plot(resid(dspi_model_ulva) ~ fitted(dspi_model_ulva))
qqnorm(resid(dspi_model_ulva))
qqline(resid(dspi_model_ulva))
```

Check the performance of the model, plot effects, get R2, make table
```{r}
performance::check_model(dspi_model_ulva)
r.squaredGLMM(dspi_model_ulva)
summary(dspi_model_ulva)
plot(allEffects(dspi_model_ulva))
tab_model(dspi_model_ulva, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
ulva_dspi_treatment_null <- lmer(formula = carbon_mol ~ Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
ulva_dspi_model2 <- lmer(formula = carbon_mol ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
anova(ulva_dspi_treatment_null, ulva_dspi_model2)
ulva_dspi_temperature_null <- lmer(formula = carbon_mol ~ Treatment + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
ulva_dspi_model3 <- lmer(formula = carbon_mol ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
anova(ulva_dspi_temperature_null, ulva_dspi_model3)
```

Plots
```{r}
ulva %>% ggplot(aes(treatment_graph, carbon_mol)) + 
        geom_boxplot(size=0.5) + 
        geom_point(alpha = 0.75, size = 3, aes(color = Temperature), show.legend = TRUE) + 
        labs(x="Treatment", y= "DSPI (mol m-2)", title= "E", subtitle = "Ulva lactuca") + 
        scale_x_discrete(labels = c("35ppt/0.5umolN", "35ppt/14umolN", "28ppt/27umolN", 
                                    "18ppt/53umolN", "11ppt/80umolN")) + 
        ylim(0, 10) + stat_mean() + 
        geom_hline(yintercept=5.0, color = "red", size = 0.5, alpha = 0.5) +
        scale_color_manual(values = c("#295102", "#7CB950", "#BDE269")) +
        theme_bw() +
        theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means for DSPI
```{r}
ulva %>% group_by(Treatment) %>% summarise_at(vars(carbon_mol), list(mean = mean))
ulva %>% group_by(Run) %>% summarise_at(vars(carbon_mol), list(mean = mean))
ulva %>% group_by(Run) %>% summarise_at(vars(day_length_avg), list(mean = mean))
```

Add growth rate from other dataset to this one and subset by species
```{r}
growth_rate <- read.csv("/Users/Angela/src/work/limu/algal_growth_photosynthesis/data_input/all_runs_growth_011723.csv")
growth_rate$Species <- as.factor(growth_rate$Species)
growth_rate$treatment <- as.factor(growth_rate$treatment)
```

Make a new column for weight change (difference final from initial)
```{r}
growth_rate$growth_rate_percent <- (growth_rate$final.weight - growth_rate$Initial.weight) / growth_rate$Initial.weight * 100
```

Subset growth data
```{r}
gr_ulva <- subset(growth_rate, Species == "Ul" & treatment != 2.5)
ulva$growth_rate <- round((gr_ulva$final.weight - gr_ulva$Initial.weight) / gr_ulva$Initial.weight * 100, digits = 2)
```

Plot growth vs dspi
```{r}
ulva_growth_dspi_graph <- ggplot(ulva, aes(x=carbon_mol, y=growth_rate)) + 
        geom_point(alpha = 0.5, size = 3, aes(color = Treatment), show.legend = TRUE) + 
        geom_smooth(method = "lm", col = "black") + theme_bw() + 
        labs(title = "A", subtitle = "Ulva lactuca", x = "DSPI (mol m-2)", 
             y = "growth rate (%)") + stat_regline_equation(label.x = 5, label.y = 225) + 
        stat_cor(label.x = 5, label.y = 215) +
        theme(plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", vjust = -20, hjust = 0.05))
ulva_growth_dspi_graph
```

Plot individual regressions for each treatment vs. growth
```{r}
ulva_growth_carbon_indiv <- ggplot(ulva, aes(x=carbon_mol, y=growth_rate)) +
        geom_point(alpha = 0.5, size = 3, aes(color = Treatment), show.legend = TRUE) +
        geom_smooth(method = "lm", col = "black") + theme_bw() +
        labs(title = "A", subtitle = "Ulva lactuca", x = "DSPI (mol m-2)", 
             y = "growth rate (%)") + 
        stat_cor() +
        #stat_regline_equation() + 
        facet_wrap(vars(Treatment), scales = "free")
ulva_growth_carbon_indiv
```

HYPNEA____________________________________________________________

Subset Hypnea data from larger dataset
```{r}
hypnea <- subset(ek_irrad_data, Species == "hm" & day1_rlc_time != "11:34:05")
hypnea$treatment_graph[hypnea$Treatment == 0] <- "1) 35ppt/0.5umol"
hypnea$treatment_graph[hypnea$Treatment == 1] <- "2) 35ppt/14umol" 
hypnea$treatment_graph[hypnea$Treatment == 2] <- "3) 28ppt/27umol" 
hypnea$treatment_graph[hypnea$Treatment == 3] <- "5) 18ppt/53umol" 
hypnea$treatment_graph[hypnea$Treatment == 4] <- "6) 11ppt/80umol"
hypnea$treatment_graph[hypnea$Treatment == 2.5] <- "4) 28ppt/53umol"
```

Make a histogram and residual plots of the data for ulva
```{r}
hist(hypnea$carbon_mol, main = paste("Hypnea musciformis"), col = "#993333", labels = TRUE)

hypnea %>% ggplot(aes(carbon_mol)) +
        geom_histogram(binwidth=0.25, fill = "#993333", color = "black", size = 0.25, alpha = 0.85) +
        theme_bw()
```

Run model without interaction between the treatments and temperature - supersat_avg is in minutes
Take RLC.Order our of the random effects because causing problems of singularity. R2 is same with or without (+ (1 | RLC.Order))
```{r}
dspi_model_hypnea <- lmer(formula = carbon_mol ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID), data = hypnea)
```

Plot residuals
```{r}
hist(resid(dspi_model_hypnea))
plot(resid(dspi_model_hypnea) ~ fitted(dspi_model_hypnea))
qqnorm(resid(dspi_model_hypnea))
qqline(resid(dspi_model_hypnea))
```

Check the performance of the model, get R2, make table, plot effects
```{r}
performance::check_model(dspi_model_hypnea)
r.squaredGLMM(dspi_model_hypnea)
summary(dspi_model_hypnea)
plot(allEffects(dspi_model_hypnea))
tab_model(dspi_model_hypnea, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
hypnea_dspi_treatment_null <- lmer(formula = carbon_mol ~ Temperature + (1 | Run) + (1 | Plant.ID) , data = hypnea, REML = FALSE)
hypnea_dspi_model2 <- lmer(formula = carbon_mol ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID), data = hypnea, REML = FALSE)
anova(hypnea_dspi_treatment_null, hypnea_dspi_model2)
hypnea_dspi_temperature_null <- lmer(formula = carbon_mol ~ Treatment + (1 | Run) + (1 | Plant.ID), data = hypnea, REML = FALSE)
hypnea_dspi_model3 <- lmer(formula = carbon_mol ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID), data = hypnea, REML = FALSE)
anova(hypnea_dspi_temperature_null, hypnea_dspi_model3)
```

Plots
```{r}
hypnea %>% ggplot(aes(treatment_graph, carbon_mol)) + 
        geom_boxplot(size=0.5) + 
        geom_point(alpha = 0.75, size = 3, aes(color = Temperature), show.legend = TRUE) + 
        labs(x="Treatment", y= "DSPI (mol m-2)", title= "F", subtitle = "Hypnea musciformis") + 
        scale_x_discrete(labels = c("35ppt/0.5umolN", "35ppt/14umolN", "28ppt/27umolN", "28ppt/53umolN", 
                                    "18ppt/53umolN", "11ppt/80umolN")) + 
        ylim(0, 10) + stat_mean() + 
        geom_hline(yintercept=5.0, color = "red", size = 0.5, alpha = 0.5) +
        scale_color_manual(values = c("#9C0627", "#BB589F", "#F4B4E2")) +
        theme_bw() +
        theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", size = 14, vjust = -20, hjust = 0.05))
```

Summarize the means for DSPI
```{r}
hypnea %>% group_by(Treatment) %>% summarise_at(vars(carbon_mol), list(mean = mean))
hypnea %>% group_by(Run) %>% summarise_at(vars(carbon_mol), list(mean = mean))
hypnea %>% group_by(Run) %>% summarise_at(vars(day_length_avg), list(mean = mean))
```

For Hypnea remove hm6-4 on 11/12 that had no d9 RLC (final weight 0.1017)
and hm6-4 on 10/29/21 because it was white and also looked dead 
```{r}
gr_hypnea <- subset(growth_rate, Species == "Hm" & final.weight != 0.1017 & growth_rate_percent > -87.96837)
hypnea$growth_rate <- round((gr_hypnea$final.weight - gr_hypnea$Initial.weight) / gr_hypnea$Initial.weight * 100, digits = 2)
```

Plot a regression between the photosynthetic independent variables of interest and growth rate
```{r}
hypnea_growth_carbon_graph <- ggplot(hypnea, aes(x=carbon_mol, y=growth_rate)) + 
        geom_point(alpha = 0.5, size = 3, aes(color = Treatment), show.legend = TRUE) + 
        geom_smooth(method = "lm", col = "black") + theme_bw() + 
        xlim(2, 8) + ylim(-100, 190) + 
        labs(title = "B", subtitle = "Hypnea musciformis", x = "DSPI (mol m-2)", 
             y = "growth rate (%)") + 
        stat_regline_equation(aes(size = 14), label.x = 6, label.y = -75) + 
        scale_color_manual(values = c("#57330F", "#22773E", "#9970C2", "#FF3333", "#FF9933", "#99FF33")) +
        stat_cor(aes(size = 14), label.x = 6, label.y = -85) +
        theme(plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05, size = 14), 
              plot.subtitle = element_text(face = "italic", vjust = -20, hjust = 0.05, size = 14))
hypnea_growth_carbon_graph
```

Plot individual regressions for each treatment vs. growth
```{r}
hypnea_growth_carbon_indiv <- ggplot(hypnea, aes(x=carbon_mol, y=growth_rate)) +
        geom_point(alpha = 0.5, size = 3, aes(color = Treatment), show.legend = TRUE) +
        geom_smooth(method = "lm", col = "black") + theme_bw() +
        labs(title = "B", subtitle = "Hypnea musciformis", x = "DSPI (mol m-2)", 
             y = "growth rate (%)") + 
        stat_cor() +
        #stat_regline_equation() + 
        facet_wrap(vars(Treatment), scales = "free")
hypnea_growth_carbon_indiv
```
