---
title: "Relative Hsat Model"
author: "Angela Richards Donà"
date: "`r Sys.Date()`"
output: 
  html_document: 
    fig_width: 9
    fig_height: 7
---

This script is similar to the ek_irrad_model but uses relative Hsat for analysis
The amount of time in Hsat was divided by the total day length, which is defined by
the amount of time when umols photons m-2 s-1 were greater than or equal to 1.

```{r, message=FALSE, warning=FALSE}
#load the various libraries
library(lme4)
library(lmerTest)
library(effects)
library(car)
library(MuMIn)
library (dplyr)
library(emmeans)
library(DHARMa)
library(performance)
library(patchwork)
library(rstatix)
#for plots and tables
library(ggplot2)
library(ggpubr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(sjPlot)
library(sjmisc)
library(mmtable2)
library(gt)
library(purrr)
library(stringr)
library(tidyr)
```

Load the dataset
```{r}
ek_irrad_data <- read.csv("input_data/mean_supersaturation_by_period.csv")
```

Assign various variables as factors
```{r}
ek_irrad_data$Run <- as.factor(ek_irrad_data$Run)
ek_irrad_data$Temperature <- as.factor(ek_irrad_data$Temp...C.)
ek_irrad_data$Treatment <- as.factor(as.character(ek_irrad_data$Treatment))
#ek_irrad_data$deltaNPQ <- as.factor(ek_irrad_data$deltaNPQ)
```

ULVA____________________________________________________________

Subset the data by species and make new column of the treatments for the plots
```{r}
ulva <- subset(ek_irrad_data, Species == "ul" & Treatment != 2.5)
ulva$treatment_graph[ulva$Treatment == 0] <- "1) 35ppt/0.5umol"
ulva$treatment_graph[ulva$Treatment == 1] <- "2) 35ppt/14umol" 
ulva$treatment_graph[ulva$Treatment == 2] <- "3) 28ppt/27umol" 
ulva$treatment_graph[ulva$Treatment == 3] <- "5) 18ppt/53umol" 
ulva$treatment_graph[ulva$Treatment == 4] <- "6) 11ppt/80umol"
ulva$treatment_graph[ulva$Treatment == 2.5] <- "4) 28ppt/53umol"
```

Make a histogram for ulva
```{r}
hist(ulva$supersat_rel, main = paste("Ulva lactuca"), col = "olivedrab3", labels = TRUE)

ulva %>% ggplot(aes(supersat_avg)) +
        geom_histogram(binwidth=5, fill = "#5BB300", color = "black", size = 0.25, alpha = 0.85) +
        theme_bw()
```

Run model without interaction between the treatments and temperature
Take RLC.Order out of the random effects because causing problems of singularity. R2 is same with or without (+ (1 | RLC.Order))
```{r}
rel_hsat_model_ulva <- lmer(formula = supersat_rel ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva)
```

Make residual plots of the data
```{r}
hist(resid(rel_hsat_model_ulva))
plot(resid(rel_hsat_model_ulva) ~ fitted(rel_hsat_model_ulva))
qqnorm(resid(rel_hsat_model_ulva))
qqline(resid(rel_hsat_model_ulva))
```

Check the performance of the model, get R2, plot the effects of model and make table
```{r fig.height=9, fig.width=7.5}
performance::check_model(rel_hsat_model_ulva)
r.squaredGLMM(rel_hsat_model_ulva)
summary(rel_hsat_model_ulva)
plot(allEffects(rel_hsat_model_ulva))
tab_model(rel_hsat_model_ulva, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
```

Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
ulva_relhsat_treatment_null <- lmer(formula = supersat_rel ~ Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
ulva_relhsat_model2 <- lmer(formula = supersat_rel ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
anova(ulva_relhsat_treatment_null, ulva_relhsat_model2)
ulva_relhsat_temperature_null <- lmer(formula = supersat_rel ~ Treatment + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
ulva_relhsat_model3 <- lmer(formula = supersat_rel ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = ulva, REML = FALSE)
anova(ulva_relhsat_temperature_null, ulva_relhsat_model3)
```

Plots
```{r}
ulva %>% ggplot(aes(treatment_graph, supersat_rel)) + 
        geom_boxplot(size=0.5) + 
        geom_point(alpha = 0.75, size = 3, aes(color = Temperature), show.legend = TRUE) + 
        labs(x="Treatment (salinty/nitrate)", y= "Relative Hsat (%)", title= "C", subtitle = "Ulva lactuca") + 
        scale_x_discrete(labels = c("35ppt/0.5umolN", "35ppt/14umolN", "28ppt/27umolN", "18ppt/53umolN", "11ppt/80umolN")) + 
        ylim(20, 100) + stat_mean() + 
        geom_hline(yintercept=60, color = "red", size = 0.5, alpha = 0.5) +
        scale_color_manual(values = c("#295102", "#7CB950", "#BDE269")) +
        theme_bw() +
        theme(legend.position = c(0.90,0.90), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", size = 14, vjust = -18, hjust = 0.05))
```

Summarize the means for relative Hsat
```{r}
ulva %>% group_by(Treatment) %>% summarise_at(vars(supersat_rel), list(mean = mean))
ulva %>% group_by(Run) %>% summarise_at(vars(supersat_rel), list(mean = mean))
ulva %>% group_by(Run) %>% summarise_at(vars(day_length_avg), list(mean = mean))
```

Add growth rate to this dataset
```{r}
growth_rate <- read.csv("/Users/Angela/src/work/limu/algal_growth_photosynthesis/data_input/all_runs_growth_011723.csv")
growth_rate$Species <- as.factor(growth_rate$Species)
growth_rate$treatment <- as.factor(growth_rate$treatment)
```

Make a new column for weight change (difference final from initial)
```{r}
growth_rate$growth_rate_percent <- (growth_rate$final.weight - growth_rate$Initial.weight) / growth_rate$Initial.weight * 100
```

Subset by species and add to current dataset
```{r}
gr_ulva <- subset(growth_rate, Species == "Ul" & treatment != 2.5)
ulva$growth_rate <- round((gr_ulva$final.weight - gr_ulva$Initial.weight) / gr_ulva$Initial.weight * 100, digits = 2)
```

Plot a regression between the photosynthetic independent variables of interest and growth rate
```{r}
ulva_growth_rel_hsat_graph <- ggplot(ulva, aes(x=supersat_rel, y=growth_rate)) + 
        geom_point(alpha = 0.5, size = 3, aes(color = Treatment), show.legend = TRUE) + 
        geom_smooth(method = "lm", col = "black") + theme_bw() + 
        labs(title = "A", subtitle = "Ulva lactuca", x = "Mean Rel. Hsat Time (%)", 
             y = "growth rate (%)") + stat_regline_equation(label.x = 75, label.y = 225) + 
        stat_cor(label.x = 75, label.y = 215) +
        theme(plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", vjust = -20, hjust = 0.05))
ulva_growth_rel_hsat_graph
```


HYPNEA_____________________________________________________________________________________________________________
Subset for Hypnea
```{r}
hypnea <- subset(ek_irrad_data, Species == "hm" & day1_rlc_time != "11:34:05")
hypnea$treatment_graph[hypnea$Treatment == 0] <- "1) 35ppt/0.5umol"
hypnea$treatment_graph[hypnea$Treatment == 1] <- "2) 35ppt/14umol" 
hypnea$treatment_graph[hypnea$Treatment == 2] <- "3) 28ppt/27umol" 
hypnea$treatment_graph[hypnea$Treatment == 3] <- "5) 18ppt/53umol" 
hypnea$treatment_graph[hypnea$Treatment == 4] <- "6) 11ppt/80umol"
hypnea$treatment_graph[hypnea$Treatment == 2.5] <- "4) 28ppt/53umol"
```

Make a histogram of the data for ulva
```{r}
hist(hypnea$supersat_rel, main = paste("Hypnea musciformis"), col = "maroon", labels = TRUE)

hypnea %>% ggplot(aes(supersat_rel)) +
        geom_histogram(binwidth=5, fill = "#a8325e", color = "black", size = 0.25, alpha = 0.85) +
        theme_bw()
```

Run model
```{r}
rel_hsat_model_hypnea  <- lmer(formula = supersat_rel ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = hypnea)
```

Plot residuals
```{r}
hist(resid(rel_hsat_model_hypnea))
plot(resid(rel_hsat_model_hypnea) ~ fitted(rel_hsat_model_hypnea))
qqnorm(resid(rel_hsat_model_hypnea))
qqline(resid(rel_hsat_model_hypnea))
```

Check the performance of the model, get R2, make table and plot effects
```{r}
performance::check_model(rel_hsat_model_hypnea)
r.squaredGLMM(rel_hsat_model_hypnea)
summary(rel_hsat_model_hypnea)
plot(allEffects(rel_hsat_model_hypnea))
tab_model(rel_hsat_model_hypnea, show.intercept = TRUE, show.se = TRUE, show.stat = TRUE, show.df = TRUE, show.zeroinf = TRUE)
```
Construct null model to perform likelihood ratio test REML must be FALSE
```{r}
hypnea_relhsat_treatment_null <- lmer(formula = supersat_rel ~ Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = hypnea, REML = FALSE)
hypnea_relhsat_model2 <- lmer(formula = supersat_rel ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = hypnea, REML = FALSE)
anova(hypnea_relhsat_treatment_null, hypnea_relhsat_model2)
hypnea_relhsat_temperature_null <- lmer(formula = supersat_rel ~ Treatment + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = hypnea, REML = FALSE)
hypnea_relhsat_model3 <- lmer(formula = supersat_rel ~ Treatment + Temperature + (1 | Run) + (1 | Plant.ID) + (1 | RLC.Order), data = hypnea, REML = FALSE)
anova(hypnea_relhsat_temperature_null, hypnea_relhsat_model3)
```

Plots
```{r}
hypnea %>% ggplot(aes(treatment_graph, supersat_rel)) + 
        geom_boxplot(size=0.5) + 
        geom_point(alpha = 0.75, size = 3, aes(color = Temperature), show.legend = TRUE) + 
        labs(x="Treatment (salinty/nitrate)", y= "Relative Hsat (%)", title= "D", subtitle = "Hypnea musciformis") + 
        scale_x_discrete(labels = c("35ppt/0.5umolN", "35ppt/14umolN", "28ppt/27umolN", "28ppt/53umolN", "18ppt/53umolN", "11ppt/80umolN")) + 
        ylim(20, 100) + stat_mean() + 
        geom_hline(yintercept=60, color = "red", size = 0.5, alpha = 0.5) +
        scale_color_manual(values = c("#9C0627", "#BB589F", "#F4B4E2")) +
        theme_bw() +
        theme(legend.position = c(0.88,0.88), plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", size = 14, vjust = -18, hjust = 0.05))
```

For regression with growth
Hypnea remove hm6-4 on 11/12 that had no d9 RLC (final weight 0.1017)
and hm6-4 on 10/29/21 because it was white and also looked dead 
```{r}
gr_hypnea <- subset(growth_rate, Species == "Hm" & final.weight != 0.1017 & growth_rate_percent > -87.96837)
hypnea$growth_rate <- round((gr_hypnea$final.weight - gr_hypnea$Initial.weight) / gr_hypnea$Initial.weight * 100, digits = 2)
```

Plot a regression between the photosynthetic independent variables of interest and growth rate
```{r}
hypnea_growth_rel_hsat_graph <- ggplot(hypnea, aes(x=supersat_rel, y=growth_rate)) + 
        geom_point(alpha = 0.5, size = 3, aes(color = Treatment), show.legend = TRUE) + 
        geom_smooth(method = "lm", col = "black") + theme_bw() + 
        labs(title = "B", subtitle = "Hypnea musciformis", x = "Mean Rel. Hsat (%)", y = "growth rate (%)") + 
        stat_regline_equation(label.x = 65, label.y = 200) + 
        stat_cor(label.x = 65, label.y = 215) +
        theme(plot.title = element_text(face = "bold", vjust = -15, hjust = 0.05), 
              plot.subtitle = element_text(face = "italic", vjust = -20, hjust = 0.05))
hypnea_growth_rel_hsat_graph
```

Summarize the means for relative Hsat
```{r}
hypnea %>% group_by(Treatment) %>% summarise_at(vars(supersat_rel), list(mean = mean))
hypnea %>% group_by(Run) %>% summarise_at(vars(supersat_rel), list(mean = mean))
hypnea %>% group_by(Run) %>% summarise_at(vars(day_length_avg), list(mean = mean))
```